---
output: html_document
---
R programming - Vectors in R
========================================================
<i>Author : Abhinav Agrawal</i>


Vector is a basic object in R.  It is a sequence of elements of the same class type.  We can have a logical vector,  numeric vector,  character vector, complex vector or a raw vector.


<b>Creating a simple vector object.</b>


Assign 1 to an object called y using the following expression :

```{r}
y = 1
```
Notice that this is just an assignment. The values of object y will not be printed on the console.

We can print the vector y by simply typing y or using the print() function.

```{r}
y

```
Here y is a numeric vector with one element and that element is 1

```{r}
print(y)
```

To check the class type of the vector object y, use the class() function on object y
```{r}
class(y)
```
Yes, we just created a very simple numeric vector, congratulations !!




<b> Vector objects with sequence of elements</b>

We can generate a sequence of elements by using <b><i>colon (:)</i></b>

```{r}
x <- 1:10
x
```

```{r}
x2 <- -1:3
x2
```




We can also generate a sequence of elements by using <b><i>c() function</i></b>, concatenate function

```{r}
xyz <- c(1,2,3,4:9)
xyz
```

We can check the class of the vector object xyz by using the class() function :

```{r}
class(xyz)
```

We can find the number of elements in the vector by using the <b><i>length()</i></b> function :

```{r}
length(xyz)
```
This shows that the vector object xyz contains 9 numeric elements.

Next, an example of a vector object of <b><i>character class</i></b> type :
```{r}
yy <- c('a','b','c')
yy
```


```{r}
class(yy)
```



Next, an example of a vector object of <b><i>complex class</b></i> type :

```{r}
zz <- c(0+1i,8+2i)
zz
```

```{r}
class(zz)
```

Another way to generate the sequence is by using the <b><i>seq() function </i></b>

```{r}
s <- seq(from = 2, to = 20, by = 3)
s
```

Another way to generate sequence is by using <b><i>rep()</i></b>, repeat function :

```{r}
rep(5,7)
```

```{r}
rep("a",4)
```

```{r}
rep(c("a","b"),2)
```

```{r}
rep(c("a","b"), each = 2)
```

```{r}
v <- vector("numeric", length =  10)
v
```